---
title: "Bayesian data analysis assignment"
output: html_notebook
---

<!-- Please don't mess with the next few lines! -->

```{=html}
<style>h5{font-size:2em;color:#0000FF}h6{font-size:1.5em;color:#0000FF}div.answer{margin-left:5%;border:1px solid #0000FF;border-left-width:10px;padding:25px} div.summary{background-color:rgba(30,144,255,0.1);border:3px double #0000FF;padding:25px}</style>
```
<!-- Please don't mess with the previous few lines! -->

## Preliminaries

```{r, message = FALSE, warning = FALSE}
library(tidyverse)
library(MASS)
library(broom)
```

##### Question 1

Two-thirds of sushi restaurants are legit while one-third are sketchy. Even legit sushi restaurants serve bad sushi 10% of the time while even sketchy sushi restaurants serve good sushi 50% of the time. Suppose you try a new sushi restaurant not knowing whether it is legit or sketchy. If the sushi is good then what is the probability you are in a legit sushi restaurant?

::: answer
| $\theta$ | $p(\theta)$ | $p(x \mid \theta)$ | $p(\theta) p(x \mid \theta)$ | $p(\theta \mid x)$  |
|----------|-------------|--------------------|------------------------------|---------------------|
| legit    | 2/3         | 9/10               | 2/3 * 9/10 = 3/5             | 3/5 / 23/30 = 18/23 |
| skechy   | 1/3         | 1/2                | 1/3 * 1/2 = 1/6              | 1/6 / 23/30 = 5/23  |
:::

##### Question 2

Solve Problem 3 from the `1_Discrete_Bayes_example` notes a different way: this time, do it sequentially by calculating the posterior probability from the first draw of a red ball and using that posterior as the prior for drawing the second red ball. (This is similar to the second solution of Problem 2 from those notes, but even easier. Since the sampling is done *with replacement*, the draws are independent, so the second draw is still out of 5 balls and the likelihood won't change from the first draw to the second.)

Your answer should be in the form of two tables like the ones in `1_Discrete_Bayes_example`, one for the first red draw and one for the second. You can copy, paste, and modify the markdown to create those tables.

::: answer
Write your answer here in table format: 1st draw

| $\theta$ | $p(\theta)$ | $p(x \mid \theta)$ | $p(\theta) p(x \mid \theta)$ | $p(\theta \mid x)$ |
|----------|-------------|--------------------|------------------------------|--------------------|
| H1       | 1/6         | 0                  | 0                            | 0                  |
| H2       | 1/6         | 1/5                | 1/6 * 1/5 = 1/30             | 1/30 / 1/2 = 1/15  |
| H3       | 1/6         | 2/5                | 1/6 * 2/5 = 1/15             | 1/15 / 1/2 = 2/15  |
| H4       | 1/6         | 3/5                | 1/6 * 3/5 = 1/10             | 1/10 / 1/2 = 1/5   |
| H5       | 1/6         | 4/5                | 1/6 * 4/5 = 2/15             | 2/15 / 1/2 = 4/15  |
| H6       | 1/6         | 1                  | 1/6 * 1 = 1/6                | 1/6 / 1/2 = 1/3    |

Write your answer here in table format: 2nd draw

| $\theta$ | $p(\theta)$ | $p(x \mid \theta)$ | $p(\theta) p(x \mid \theta)$ | $p(\theta \mid x)$    |
|----------|-------------|--------------------|------------------------------|-----------------------|
| H1       | 0           | 0                  | 0                            | 0                     |
| H2       | 1/15        | 1/5                | 1/15 * 1/5 = 1/75            | 1/75 / 11/15 = 1/55   |
| H3       | 2/15        | 2/5                | 2/15 * 2/5 = 4/75            | 4/75 / 11/15 = 4/55   |
| H4       | 1/5         | 3/5                | 1/5 * 3/5 = 9/75             | 9/75 / 11/15 = 9/55   |
| H5       | 4/15        | 4/5                | 4/15 * 4/5 = 16/75           | 16/75 / 11/15 = 16/55 |
| H6       | 1/3         | 1                  | 1/3 * 1 = 25/75 = 1/3        | 1/3 / 11/15 = 5/11    |
:::

##### Question 3

Suppose that we wish to predict whether a given stock will issue a dividend this year ("Yes" or "No") based on $x$, last year's profit (in some unspecified units that don't matter for this problem). We examine a large number of companies and discover that the mean value of $x$ for companies that issued a dividend was $\bar{x} = 10$, while the mean for those that didn't was $\bar{x} = 0$. In addition, the standard deviation of $x$ for these two sets of companies was $\sigma = 6$. Finally, 80% of companies issued dividends. Assuming that $x$ follows a normal distribution, **predict the probability that a company will issue a dividend this year given that its profit was** $x = 4$ last year.

We'll use Bayes's Theorem to solve this problem. The parameter $\theta$ is a discrete parameter that can be either $Y$ or $N$ (for "Yes" or "No") representing whether a stock issues a dividend or not.

Since the process is a little involved, we'll break it down into easy-to-digest chunks. Make sure you're clear on each step before moving onto the next.

###### (a).

Write down a contextually meaningful interpretation of the mathematical expression $p(\theta = Y \mid x = 4)$. (This is the posterior that we need to calculate.)

::: answer
The probability that a company will issue a dividend this year given that the profit was 4 units.
:::

###### (b).

Write down a contextually meaningful interpretation of the prior $p(\theta = Y)$.

::: answer
The prior that a randomly chosen company will issue a dividend this year.
:::

###### (c).

According to the information given in the problem statement, what is the value of the prior $p(\theta = Y)$? Store this value in a variable called `prior`.

::: answer
```{r}
# Uncomment the lines below and replace the question mark to
# store the value of the prior in a variable called prior.

prior <- .8
prior
```
:::

###### (d).

Write down a contextually meaningful interpretation of the likelihood $p(x = 4 \mid \theta = Y)$.

::: answer
The probability that a company's profit was 4 units given that they gave out dividends.
:::

###### (e).

To calculate the likelihood $p(x = 4 \mid \theta = Y)$, we recall that last year's profits for companies that issue dividends are distributed according to the normal model $N(10, 6)$. The general formula for a normal model is

$$\frac{1}{\sigma \sqrt{ 2 \pi}}e^{-(x - \mu)^{2}/2\sigma^{2}}$$ Plug in all the known values to obtain a value for the likelihood. The R command for the square root is `sqrt`. The number $\pi$ is stored as `pi`. Powers are computed using the caret `^` except the exponential function, which uses `exp`. For example, $e^{2\sqrt{1 + \pi^{2}}}$ would be calculated using the R code `exp(2 * sqrt(1 + pi^2))`. Be careful with parentheses and order of operations!

Store your answer in a variable called `likelihood`. (Hint: the answer should be 0.04032845)

::: answer
```{r}
# Uncomment the lines below and replace the question mark to
# evaluate the normal model N(10, 6) at x = 4
# and store the value in a variable called likelihood.

x <- 4.0
mu <- 10.0
sigma <- 6.0

likelihood = (1 / (sigma * sqrt(2 * pi))) * exp(-((x - mu)^2) / (2 * sigma^2))
likelihood
```
:::

###### (f).

The denominator of Bayes's Theorem (the "evidence") is

$$\sum_{\theta} p(\theta) p(x \mid \theta).$$

In this problem, the sum is only two terms, corresponding to the two possible values of $\theta$ ($Y$ or $N$). So it looks like the following:

$$p(\theta = Y) p(x = 4 \mid \theta = Y) + p(\theta = N) p(x = 4 \mid \theta = N).$$

You already know the first term of the sum $p(\theta = Y) p(x = 4 \mid \theta = Y)$. (This is just `prior * likelihood`.)

What is the value of $p(\theta = N)$? (You know $p(\theta = Y)$, so don't overthink this!)

::: answer
$$p(θ=N) = 1−0.8 = 0.2$$
:::

###### (g).

To compute $p(x = 4 \mid \theta = N)$?, we'll do something very similar to what we did to compute the likelihood in Part (e). But now the normal model changes from $N(10, 6)$ to $N(0, 6)$. Why?

::: answer
$$p(x = 4 \mid \theta = N) = \frac{1}{6 \sqrt{ 2 \pi}}e^{\frac{-(4 - 0)^{2}}{2*6^{2}}}$$
:::

###### (h).

Using the normal model $N(0, 6)$, compute $p(x = 4 \mid \theta = N)$.

::: answer
```{r}
# Evaluate the normal model N(0, 6) at x = 4.

p_no <- 0.2
mu_no <- 0.0

likelihood_no <- (1 / (sigma * sqrt(2 * pi))) * exp(-((x - mu_no)^2) / (2 * sigma^2))

p_no
likelihood_no
```
:::

###### (i).

Put parts (f), (g), and (h) all together now to compute the denominator for Bayes's Theorem:

$$p(\theta = Y) p(x = 4 \mid \theta = Y) + p(\theta = N) p(x = 4 \mid \theta = N).$$

Store your answer in a variable called `evidence`.

::: answer
```{r}
# Uncomment the lines below and replace the question mark to
# evaluate the denominator for Bayes's Theorem (the evidence)
# and store the value in a variable called evidence.

evidence <- prior * likelihood + p_no * likelihood_no
evidence
```
:::

###### (j).

If everything worked correctly, the following code should compute the correct posterior probability:

```{r}
# Uncomment the lines below to calculate and report the posterior.

posterior <- prior * likelihood / evidence
posterior
```

The answer should be 0.7518525. (If you didn't get that, go back and check your work *carefully*!) State this result as a conclusion in the context of the original problem.

::: answer
Given that the company's profit last year was 4 units, the probability that the company issued a dividend was about 75.185%
:::
